Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-03 demystify.tex
Updates from Overleaf

TA is common in the beginning when you haven't found an adviser who would pay you RA. As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation). Your TA is paid through the school or department, i.e., they hire you to help teach.  During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding is not typically available during the summer, which has few courses.

\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for good CS PhD programs. When you apply to be a full-time student,  state that you need financial assistant. It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.

@@

\textbf{How to get RA?} When a professor recruits you, they might offer you an RA right away (e.g., when you apply).  A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA. See how to contact a prof. in \autoref{sec:contact} for research opportunities.

@@

For example, in the 2024 admission cycle, there is a huge surge in students who are interested in AI and NLP (thanks, ChatGPT!). However, our AI/NLP faculty cannot take more students and therefore we cannot admit many applicants, many of whom have excellent profiles (e.g., papers in top conferences, strong LoRs, etc.).

>>>>

TA is common in the beginning when you haven't found an adviser who would pay you RA. It is also common to sandwich between TA and RA (e.g., when your prof. does not have sufficient funding or you want to try get TA experience).

Your TA is paid through the school or department, i.e., they hire you to help teach.
As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation).
During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding is not typically available during the summer, which has few courses.

\paragraph{How to get TA?}  Unless you have other funding such as RA or Fellowships, TA is typically default for CS PhD programs. When you apply to be a full-time student,  state that you need financial assistant. It is common that the PhD committee will either admit you and give you GTA, or reject you; i.e., we do not admit a student without supporting them.

@@

\textbf{How to get RA?} When a professor recruits you, they might offer you an RA right away (e.g., when you apply).  A common scenario is that you first get admitted with TA, and then after a year or two find an adviser to support you with RA. RA is never guaranteed as it depends on the funding situation of the prof. So you should also pay attention to TA, e.g., see if TA is readily available for PhD students in the program.

@@

For example, in the 2024 admission cycle, there is a huge surge in students interested in AI and NLP (thanks, ChatGPT!). Thus, AI/NLP faculty might be overwhelm with potential students and cannot take additional students, many of whom have excellent profiles.
